---
title: "SIMD 2020 Domain Explorer"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: menu
    source_code: embed
    self_contained: false
    theme:
      version: 4
      bootswatch: journal
---

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(DT)
library(scales)

sf_datazones <- read_sf("Shapefile/cleaned.shp")

theme_set(theme_classic())
theme_update(panel.grid.major.y = element_line(),
             panel.grid.major.x = element_line(),
             plot.title.position = "plot")
```
  

Overall SIMD2020
=====================================
  
Column 
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "SIMD Decile: ", sf_datazones$rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "SIMD 2020\nDecile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------


Income
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$inc_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Income Decile: ", sf_datazones$inc_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(inc_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Income Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = inc_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(1,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(1,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Income Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Employment
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$emp_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Employment Decile: ", sf_datazones$emp_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(emp_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Employment Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = emp_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Employment Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")

```


Health
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$hlth_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Health Decile: ", sf_datazones$hlth_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(hlth_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Health Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = hlth_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Health Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Education
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$edu_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Education Decile: ", sf_datazones$edu_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(edu_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Education Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = edu_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Education Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Geographic Access
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$g_acc_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Access Decile: ", sf_datazones$g_acc_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(g_acc_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Access Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = g_acc_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Geographic Access Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Crime
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$crime_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Crime Decile: ", sf_datazones$crime_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(crime_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Crime Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = crime_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Crime Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Housing
=====================================
  
Column {data-width=600}
-------------------------------------

### Map

```{r}
# Palette
mypalette <- colorFactor(palette = "viridis", domain = sf_datazones$house_rank_decile, na.color = "transparent")

# Text
mylabels <- paste(
  "Datazone: ", sf_datazones$dz_name,"<br/>",
  "Council Area: ", sf_datazones$la_name,"<br/>",
  "Housing Decile: ", sf_datazones$house_rank_decile,
  sep = "") %>% 
  lapply(htmltools::HTML)

# Plot
m <- leaflet(sf_datazones) %>% 
  addTiles() %>% 
  setView( lng = -4.182,
           lat = 56.817,
           zoom = 7) %>% 
  addPolygons(
    fillColor = ~mypalette(house_rank_decile),
    weight = .5,
    opacity = 1,
    color = "grey",
    fillOpacity = .5,
    label = mylabels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) %>% 
  addLegend(
    pal = mypalette,
    values = ~rank_decile,
    opacity = .7,
    title = "Housing Decile",
    position = "bottomright"
  )

m
```
   
Column {data-width=400}
-------------------------------------

### Rank Comparison

```{r}
ggplot(data = sf_datazones, aes(x = rank, y = house_rank)) +
  geom_point(size = 1.5, alpha = .5, colour = "dodgerblue4") +
  scale_x_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,by=1000), labels = comma) +
  labs(x = "Overall Rank", y = "Housing Rank", subtitle = "Areas are ranked from 1 (Most Deprived) to 6,976 (Least Deprived)")
```


Data
=====================================

Column
-------------------------------------

### Data {data-height=900}

```{r}
sf_datazones %>% 
  st_drop_geometry() %>% 
  select(dz_name, la_name, rank_decile, inc_rank_decile, emp_rank_decile, hlth_rank_decile, edu_rank_decile, g_acc_rank_decile, crime_rank_decile, house_rank_decile) %>%
  rename("Datazone" = "dz_name",
         "Council Area" = "la_name",
         "Income Decile" = "inc_rank_decile",
         "Employment Decile" = "emp_rank_decile",
         "Health Decile" = "hlth_rank_decile",
         "Education Decile" = "edu_rank_decile",
         "Access Decile" = "g_acc_rank_decile",
         "Crime Decile" = "crime_rank_decile",
         "Housing Decile" = "house_rank_decile",
         "Overall Decile" = "rank_decile") %>% 
  DT::datatable(rownames = F, options = list(pageLength = 100))
```   

### Sources

Information on the Scottish Index of Multiple Deprivation 2020 (v2) is available here: https://www.gov.scot/collections/scottish-index-of-multiple-deprivation-2020/

SIMD2020 Shapefiles are available here: https://data.gov.uk/dataset/1102bf85-ed49-440a-b211-da87e8d752eb/scottish-index-of-multiple-deprivation-simd-2020

You can find the files used to create this site here: https://github.com/will-ball/simd_domain_explorer